---
title: "The IPBES long-term vision of data and knowledge management"
subtitle: "Version {{< meta version >}} · {{< meta release_date >}}"
version: "1.1"
release_date: "April 2022"
concept_doi: "10.5281/zenodo.5513473"
contact_email: "tsu.data@ipbes.net"
release_notes: |
  - This is a test release
  - Initial living-document baseline established.
  - YAML-driven track generation removed in favor of a single-source document.
  - Rendering switched to Quarto website defaults with `_site` output.
editors:
  - name: "Rainer M Krug"
    institution: "University of Zurich"
    orcid: "0000-0002-7490-0066"
  - name: "Aidin Niamir"
    institution: "Senckenberg Society for Nature Research"
    orcid: "0000-0003-4511-3407"
format:
  html:
    output-file: index.html
    toc: true
    number-sections: true
    format-links: false
  pdf:
    papersize: a4
    toc: true
    number-sections: true
    geometry:
      - left=2.0cm
      - right=2.5cm
      - top=1.0cm
      - bottom=1.4cm
      - includeheadfoot
    classoption:
      - titlepage
    template-partials:
      - templates/title.tex
execute:
  echo: false
---

```{r}
`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0 || identical(x, "")) y else x
}

meta <- rmarkdown::metadata
authoritative_title <- as.character(meta$title %||% "IPBES long-term vision on data and knowledge management")
authoritative_version <- as.character(meta$version %||% "development")
authoritative_release_date <- as.character(meta$release_date %||% as.character(Sys.Date()))
env_doi <- trimws(Sys.getenv("IPBES_DOI", unset = ""))
authoritative_doi <- if (!identical(env_doi, "")) env_doi else "Will be inserted upon release"
authoritative_concept_doi <- as.character(meta$concept_doi %||% "")
authoritative_release_notes <- as.character(meta$release_notes %||% "")
meta_editors <- meta$editors

title_text <- authoritative_title
version <- authoritative_version
release_date <- substr(authoritative_release_date, 1, 10)
doi <- authoritative_doi
concept_doi <- authoritative_concept_doi
doi_url <- ""
concept_doi_url <- ""
release_notes <- Sys.getenv("IPBES_RELEASE_NOTES", unset = authoritative_release_notes)
build_pdf <- tolower(Sys.getenv("IPBES_BUILD_PDF", unset = "false")) %in% c("1", "true", "yes", "y", "on")
doi_missing <- identical(trimws(doi), "") || identical(tolower(trimws(doi)), "will be inserted upon release")
if (identical(doi_url, "") && !doi_missing) {
  doi_url <- paste0("https://doi.org/", doi)
}
if (identical(concept_doi_url, "") && !identical(concept_doi, "")) {
  concept_doi_url <- paste0("https://doi.org/", concept_doi)
}
editor_names <- character(0)
if (is.list(meta_editors) && length(meta_editors) > 0) {
  editor_names <- vapply(meta_editors, function(x) as.character(x$name %||% ""), character(1))
  editor_names <- editor_names[nzchar(editor_names)]
}
editor_text <- if (length(editor_names) > 0) {
  paste(editor_names, collapse = ", ")
} else {
  "Editors TBD"
}
release_year <- substr(release_date, 1, 4)
citation_line <- if (!doi_missing && !identical(doi_url, "")) {
  sprintf(
    "IPBES (%s): *%s*, version %s, %s (eds.). IPBES secretariat, Bonn, Germany. DOI: <%s>",
    release_year, title_text, version, editor_text, doi_url
  )
} else {
  sprintf(
    "IPBES (%s): *%s*, version %s, %s (eds.). IPBES secretariat, Bonn, Germany. DOI: **Will be inserted upon release**",
    release_year, title_text, version, editor_text
  )
}
pdf_line <- if (build_pdf) {
  "<a href=\"assets/IPBES-Data-Management-Vision-latest.pdf\" target=\"_blank\" rel=\"noopener noreferrer\">Download latest PDF</a>"
} else {
  "Generated in release/CI builds."
}
```

\newpage

## Release Metadata {.unnumbered}

- **Version:** `r version`
- **Release date:** `r release_date`
- **DOI:** `r if (!doi_missing && !identical(doi_url, "")) sprintf("[%s](%s)", doi, doi_url) else "**Will be inserted upon release**"`
- **Latest-version DOI:** `r if (!identical(concept_doi, "") && !identical(concept_doi_url, "")) sprintf("[%s](%s)", concept_doi, concept_doi_url) else "N/A"`
- **PDF:** `r if (knitr::is_html_output()) pdf_line else "Download latest PDF from the website."`

## Release Notes {.unnumbered}

```{r, results='asis'}
if (!identical(trimws(release_notes), "")) {
  cat(release_notes)
} else {
  cat("N/A")
}
```

```{r, results='asis'}
if (knitr::is_html_output()) {
  cat(
    "<p>",
    "<a href=\"index.html\">Vision</a>",
    " | ",
    "<a href=\"older-releases.html\">All Releases</a>",
    " | ",
    "<a href=\"assets/IPBES-Data-Management-Vision-latest.pdf\" target=\"_blank\" rel=\"noopener noreferrer\">Latest PDF</a>",
    "</p>",
    sep = ""
  )
}
```


\newpage

## Introduction

The long-term vision on data and knowledge management provides the
overall direction for how IPBES will organise, manage and deliver data
and knowledge across its work programme up to 2030. It sets out the
long-term ambition of the Platform in this area and outlines the
targets and pathways needed to achieve it. While the data and
knowledge management policy defines the principles, standards and
procedures that govern data and knowledge management within IPBES, the
long-term vision describes the desired state toward which these
efforts are directed.

This vision outlines the main objectives, guiding principles and
operational tracks that will ensure that all IPBES products are
transparent, traceable, accessible and reusable. It supports the
delivery of the four core functions of IPBES and complements both the
data and knowledge management policy and the code of practice on the
use of artificial intelligence, tailored to the needs of assessments.
It also provides the framework for the continued integration of the
FAIR and CARE principles in all aspects of data and knowledge
management within IPBES.

The implementation of this vision will be pursued through two closely
linked tracks: (i) data and knowledge management within IPBES,
focusing on the organisation, documentation and curation of
information used and generated in assessments; and (ii) delivery and
monitoring of IPBES products, focusing on their accessibility,
interoperability and long-term preservation. Together, these tracks
form a coherent structure to guide actions up to 2030 and to inform
the development of a strategic roadmap in the next intersessional
period.

## Vision and aims

By 2030, IPBES ensures the open and transparent exchange of knowledge
and data between science, policy and society regarding its products
and processes, and contributes to strengthening similar exchanges more
broadly in support of biodiversity, ecosystem services and human
well-being. Data, processes and workflows supporting the Platform in
fulfilling its mandate will be fully compliant with the FAIR and CARE
principles and will make responsible use of new and emerging
technologies, including artificial intelligence, to enhance the
quality, efficiency, accessibility, and distribution of IPBES
products.

Pursuing this vision will foster stronger collaboration, greater
efficiency in production and distribution, and improved impact of
IPBES products in bridging science, policy and society. The following
aims provide the basis for moving towards this vision:

- Ensure that all IPBES products and their underlying knowledge and data
  follow the FAIR and CARE principles;

- Strengthen and institutionalise connections within IPBES concerning
  data and knowledge, including collaboration among all task forces and
  technical support units;

- Strengthen engagement with all relevant stakeholders, including
  Indigenous Peoples and local communities, on matters of data and
  knowledge, to enhance the scientific credibility, legitimacy and
  uptake of IPBES products; and

- Regularly assess new and advancing technologies, including artificial
  intelligence, for their potential to support this vision and suggest
  their use where appropriate.

## Roadmap up to 2030

The roadmap up to 2030 is structured around two complementary tracks
that together support the implementation of the long-term vision. Each
track addresses one essential aspect of data and knowledge management
and is accompanied by specific targets designed to enhance the
transparency, accessibility and interoperability of IPBES products.
Progress within these tracks will ensure the realisation of the four
aims of the long-term vision and strengthen the science policy
interface across all functions of the Platform.

## Track 1: Data and knowledge management within IPBES

- **Body of knowledge** - By 2030, IPBES will have established the IPBES
  body of knowledge (IPBES BoK), a coherent framework connecting all
  data and knowledge used or generated through IPBES assessments. The
  BoK will document workflows, variables, indicators and identified
  knowledge gaps to ensure full transparency and traceability. It will
  serve as the central reference for linking assessment findings to
  their supporting evidence and will form the backbone for the future
  delivery of IPBES products.

- **Policy framework** - All data and knowledge included in the IPBES
  BoK will follow the standards and procedures set out in the data and
  knowledge management policy and its annexes, including the FAIR and
  CARE principles and the code of practice on the use of artificial
  intelligence, tailored to the needs of assessments. The overarching
  goal is to improve data traceability, accountability, transparency,
  interoperability, reproducibility and ethical use across all IPBES
  activities.

- **Standards and identifiers** - Adherence to widely recognised
  interoperability standards will facilitate discovery, access and
  citation of data and knowledge. Persistent identifiers will be used
  for attribution and citation, ensuring that the original authors and
  contributors receive credit and that the use of IPBES products is
  traceable.

- **Access and stakeholders** - IPBES members and stakeholders,
  including Indigenous Peoples and local communities, will be able to
  access and use IPBES products and the underlying knowledge and data
  documented in their production, ensuring equitable benefits from their
  use.

- **Scope and external data** - While IPBES products rely extensively on
  external data and knowledge, including Indigenous and local knowledge,
  their management remains outside the direct scope of IPBES. However,
  IPBES will provide guidance encouraging the use of external data that
  follow the FAIR and CARE principles and other provisions of the data
  and knowledge management policy.

## Track 2: Delivery and monitoring of IPBES products

- **Data delivery platform** - By 2030, IPBES will operate a data
  delivery platform (IPBES DDP) that provides user friendly access to
  IPBES products, the data generated during their production and curated
  links to relevant external datasets. The DDP will draw on the IPBES
  BoK and follow the data and knowledge management policy to ensure
  consistent metadata, standards and attribution.

- **Monitoring and reporting** - Real time monitoring will provide
  information on the use and reach of IPBES products and on continued
  adherence to the data and knowledge management policy.

- **Connections (internal and external)** - Connections between IPBES
  products and datasets will be maintained and published to improve
  accessibility and enable cross product searches. Links between IPBES
  products and relevant external documents, such as national
  environmental policies, private sector standards and external data
  providers, will also be identified and published where appropriate.

- **Dynamic updating and versioning** - IPBES will implement tools to
  monitor the availability of underlying external datasets, check
  compliance with the data and knowledge management policy and allow the
  automatic update of figures, tables and other visualisations that rely
  on dynamically changing data. Versioning of final IPBES products will
  be introduced to enable updates based on continuously delivered data,
  aligned with the concept of living or dynamic documents.

- **Stakeholder access and ILK sharing** - The use of IPBES products,
  including the uptake of identified knowledge gaps, will be monitored
  and shared with IPBES members and stakeholders, including Indigenous
  Peoples and local communities. IPBES will ensure that final products
  incorporating Indigenous and local knowledge are shared back with the
  relevant knowledge holders to support dialogue and equitable benefit
  sharing.

- **Training and feedback** - By 2030, IPBES will offer online training
  materials demonstrating how to access, navigate and use IPBES
  products, data and associated knowledge, along with open feedback
  channels for users and stakeholders.

## Targets of the roadmap

The targets for data and knowledge management and delivery up to 2030
are aligned with the six objectives of the IPBES rolling work
programme to 2030: (1) assessing knowledge; (2) building capacity; (3)
strengthening knowledge foundations; (4) supporting policy; (5)
communicating and engaging; and (6) improving platform effectiveness.
Progress toward these targets will guide implementation of the two
tracks of the long-term vision and contribute to achieving the four
aims of the long-term vision.

## Track 1: Data and knowledge management within IPBES

Target 1.1 -- By 2023, set up a public bibliographic library
containing the underlying bibliographic records of IPBES products.
(Objectives 1 & 3)

Target 1.2 -- Starting in 2024, revise the IPBES data and knowledge
management policy every two years and support its consistent
implementation across the Platform. (Objective 3)

Target 1.3 -- By 2025, ensure that all IPBES products adhere to the
CARE principles (Collective Benefit, Authority to Control,
Responsibility and Ethics). (Objectives 3 & 5)

Target 1.4 -- By 2026, establish the IPBES BoK as a structured and
publicly findable framework documenting all underlying resources used
in IPBES products. (Objective 3)

Target 1.5 -- By 2027, ensure that all IPBES products fully adhere to
the FAIR principles (Findable, Accessible, Interoperable and
Reusable). (Objective 3)

## Track 2: Delivery and monitoring of IPBES products

Target 2.1 -- By 2026, disseminate knowledge gaps identified in IPBES
products in a structured format with persistent identifiers to enable
future monitoring and tracking. (Objective 3)

Target 2.2 By 2026, establish explicit computational links between
IPBES products and relevant policy tools and methodologies to enhance
policy support. (Objectives 3 & 4)

Target 2.3 -- By 2025, develop a roadmap to strengthen the capacities
of stakeholders for accessing and using IPBES data and knowledge.
(Objectives 2, 3 & 5)

Target 2.4 -- By 2026, develop a roadmap to strengthen the capacities
of stakeholders, including policy makers, for accessing and using
IPBES data and knowledge. (Objectives 2, 3 & 5)

Target 2.5 -- By 2027, monitor the degree to which external data
sources follow the data management and governance principles set out
in the IPBES data and knowledge management policy in order to
encourage open science. (Objective 3)

Target 2.6 -- By 2027, monitor the use of and identify potential gaps
in tools and methodologies for the delivery of IPBES products to
decision makers. (Objectives 4 & 6)

Target 2.7 -- By 2027, collect feedback from IPBES members and
stakeholders, including Indigenous Peoples and local communities and
policy makers, on the accessibility, quality and reach of IPBES
products in order to improve their impact. (Objectives 5 & 6)

Target 2.8 -- By 2030, create a technical platform capable of
dynamically updating selected aspects of IPBES products according to
the availability of new resources. (Objectives 1 & 3)

Target 2.9 -- By 2030, develop a tracking system to map the use and
citation of IPBES products in national, regional and global policies as
well as in scientific publications. (Objectives 3 & 4)

## Suggested Citation

```{r, results='asis'}
cat(citation_line)
```
